Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-15 thesis/paper_draft.qmd
refine abstract

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and lower accuracy. To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints. Among these, **MinT** (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors. The current practice is to use the shrinkage estimator (often shrinking toward a diagonal matrix), but it lacks flexibility and might not fully utilise the prominent latent structure presented. In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, namely NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance), PC-adjusted estimators (taking the latent factors directly into its construction), and others.

>>>>

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these series without respecting the hierarchical nature of the data can lead to incoherent forecasts across aggregation levels and lower accuracy. To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted to satisfy the aggregation constraints. Among these, MinT (Minimum Trace) is widely used, however, it requires a good estimate of the covariance matrix of the base forecast errors. The current practice is to use the shrinkage estimator (shrinking toward a diagonal matrix) for 1-step-ahead covariance estimation and proportionally scale it to approximate the h-step-ahead covariance matrix, but it leaves a question of whether this is appropriate for all real-world applications. We propose several alternative covariance estimators that address the shortcomings of the shrinkage estimator, including the NOVELIST estimator (which shrinks toward a soft-thresholded target), the PC-adjusted estimators (which utilises the latent factors structure), and estimators that relax the proportional scaling assumption. We evaluate the performance of MinT with these different covariance estimators in both point and probabilistic forecast reconciliation settings, and demonstrate their effectiveness and improvements over the shrinkage estimator in a complex, large-hierarchy dataset.
